Carbonate chemistry at Ny-Ålesund

### 2021-03-02 awipev-CO2.Rmd
pCO2_inst replaced by pco2_fb select -pco2_inst cutoff date: 2021-01-10

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & pressure_insitu_ctd <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & pressure_insitu_ctd <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & pressure_insitu_ctd <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 9, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
```

@@

# Monthly distribution of hourly data
d2 <- d_all %>% 
  dplyr::select(montht, pco2_inst, phINT)
count_pco2_inst_no_NA <- sum(!is.na(d2$pco2_inst)) # total non-NA pCO2 values
count_phINT_no_NA <-  sum(!is.na(d2$phINT)) # total non-NA pH seafet values
# calculate the distribution of non-NA pCO2 and pH values per month
d2_summary <- d2 %>% 
  dplyr::group_by(montht) %>% 
  dplyr::summarise(pco2_pc = round(100 * sum(!is.na(pco2_inst))/count_pco2_inst_no_NA, digits = 1),
                   pH_pc = round(100 * sum(!is.na(phINT))/count_phINT_no_NA, digits = 1)
                   )
# plot the distribution of non-NA pCO2 and pH values per month
d2_summary_long <- d2_summary %>% 
  tidyr::pivot_longer(cols = c(pco2_pc, pH_pc))
p_avail_month <- ggplot(data = d2_summary_long) +
  geom_col(aes(x = montht, y = value, fill = name), position = "dodge", width = 0.75) +
  scale_fill_discrete(name = "", labels = c("pCO2", "pH")) + 
  labs(x="Month", y="Percent of total number of data", title = "") +
  Mytheme(size_labs = 8) +
  theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = "cm"))
#ggsave(filename = "figures/monthly_availability.png", plot = p_avail_month, width = 15, height = 8, units = "cm")

@@

# ```{r whole time series, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
# size_point <- 2 # size of data points
# alpha <- 1 #transparency
# 
# d <- d_all %>% 
#   as_tsibble(index = datetime) %>% 
#   dplyr::arrange(datetime) %>%
#   tsibble::fill_gaps() %>%
#   dplyr::mutate(year_month = format(datetime, '%Y-%m'),
#                 quarter = quarter(datetime)) %>%
# #  dplyr::rename(pco2_contros = pco2_corr) %>% 
#   replace(., is.na(.), NA) # replace NaN by NA everywhere
# 
# d_ts <- d %>%
#   dplyr::select(datetime, temp_insitu_11m, sal_fb, par_air, pco2_inst,
#                 pHint_tot_sf_9m, ph_s_sf_temp_insi, at_calc, at) #%>%
#   #tidyr::pivot_longer(-datetime, values_to = "value")
#   
# #sal_fb
# ts_sal <- d_all %>%
#   ggplot(aes(x = datetime, y = sal_fb), na.rm=TRUE) +
#   scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
#   geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
#   labs(title="Salinity FerryBox",x="", y="Salinity") +
#   Mytheme(size_labs = 8)
# # ts_sal_py <- ggplotly(ts_sal,dynamicTicks = TRUE)
# # filename.html <- paste0(path_fig, "ts_sal_py.html")
# # saveWidget(ts_sal_py, filename.html, selfcontained = T, libdir = "lib")
# 
# # temp_insitu_11m
# ts_temp <- d_all %>%
#   ggplot(aes(x = datetime, y = temp_insitu_11m), na.rm=TRUE) +
#   scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
#   geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
#   labs(title="In situ temperature 11 m",x="", y="Temperature (°C)") +
#   Mytheme(size_labs = 8)
# 
# # PAR
# ts_par_air <- d_all %>%
#   ggplot(aes(x = datetime, y = par_air), na.rm=TRUE) +
#   scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
#   geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
#   labs(title="Atmospheric PAR",x="", y="PAR (umol m-2 s-1)") +
#   Mytheme(size_labs = 8)
# 
# # PCO2 
# ts_pCO2 <- d_all %>%
#   ggplot() +
#   scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
#   geom_point(aes(x = datetime, y = pco2_inst), 
#              colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
#   geom_point(aes(x = datetime, y = pco2_calc_fb), colour="red", alpha=alpha, na.rm=TRUE, size=size_point) + 
#   labs(title="Mesured (blue) and calculated (CT+AT; red) pCO2",x="", y="pCO2 (uatm)") +
#   Mytheme(size_labs = 8)
# 
# # SEAFET pH
# ts_pH <- d_all %>%
#   ggplot(aes(x = datetime, y = pHint_tot_sf_9m), na.rm=TRUE) +
#   scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
#   geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
#   geom_point(aes(x = datetime, y = ph_s_sf_temp_insi), colour="red", alpha=alpha, na.rm=TRUE, size=size_point) + 
#   labs(title="In situ pH below 8 m (seaFET in blue, spectrophotometric pH in red))",x="", y=expression(paste(pH[T]))) +
#   Mytheme(size_labs = 8)
# 
# # AT
# ts_at <- d_all %>%
#   ggplot(aes(x = datetime, y = at_calc), na.rm=TRUE) +
#   scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
#   geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
#   geom_point(aes(x = datetime, y = at), colour="red", alpha=alpha, na.rm=TRUE, size=size_point) + 
#   labs(title="Total alkalinity calculated from salinity (blue) and discrete measurements (red)", x="", y="AT (umol/kg)") +
#   Mytheme(size_labs = 8)
# 
# g <- cowplot::plot_grid(ts_sal, ts_temp, ts_par_air, ts_pCO2, ts_pH, ts_at, align='v', ncol=1)
# print(g)
# ```
## Maintenance
<div style="text-align: justify">

>>>>

```{r read observatory data, cleanup and prepare ferrybox data, echo=FALSE, warning=FALSE, message=FALSE}
d_hour <-
  readRDS(file = paste0(
    path,
    "fb_awipev-co2_server/ny-alesund/data/NRT_data/nydata_hour.rds"
  ))
d_hour <- d_hour %>%
  dplyr::filter(datetime <= as.POSIXct("2021-01-10")) %>% #cutoff date
  dplyr::select(-pco2_inst) %>% # this is the serial number of pCO2 sensor, not needed
  dplyr::rename(
    pco2_fb = pco2_corr_filtered,
    temp_dur = temp_dur_filtered,
    pco2_raw = pco2_raw_filtered,
    sal_fb = sal_fb_filtered,
    temp_insitu_11m = temp_insitu_11m_filtered,
    temp_fb = temp_fb_filtered,
    ph_dur = ph_dur_filtered
  ) %>% 
  arrange(datetime)
d_all <- as_tibble(d_hour) #d_all will be the merger
d_all <- d_all %>%
  dplyr::mutate_all(function(x) ifelse(is.nan(x), NA, x)) %>% # replace all NaN by NA
  dplyr::mutate(depth = ifelse( # create new variable depth
      pressure_insitu_ctd <= 2,
      1,
      ifelse(
        pressure_insitu_ctd > 2 & pressure_insitu_ctd <= 4,
        3,
        ifelse(
          pressure_insitu_ctd > 4 & pressure_insitu_ctd <= 6,
          5,
          ifelse(
            pressure_insitu_ctd > 6 & pressure_insitu_ctd <= 8,
            7,
            ifelse(pressure_insitu_ctd > 8, 9, NA)
          )
        )
      )
    )) %>%
  dplyr::rowwise() %>% 
  dplyr::mutate(n_depth_ctd = sum(is.na( # create new variable to identify the hours for which there are CTD data at more than 1 depth 
  c(
    sal_insitu_ctd_1m,
    sal_insitu_ctd_3m,
    sal_insitu_ctd_5m,
    sal_insitu_ctd_7m,
    sal_insitu_ctd_9m
  )
), na.rm = TRUE)) %>% 
  ungroup()
#restore datetime, date which are mangled by the mutates above, arrange ascending
d_all <- d_all %>% 
  dplyr::mutate(datetime = d_hour$datetime, 
                date = d_hour$date,
                year = lubridate::year(d_hour$datetime), # year as a number
                monthd = lubridate::month(d_hour$datetime), # month as a number
                montht = lubridate::month(d_hour$datetime, label = TRUE, abbr = TRUE) # abbreviated months
  )
```

@@

# Monthly distribution of hourly data
d2 <- d_all %>% 
  dplyr::select(montht, pco2_fb, phINT)
count_pco2_fb_no_NA <- sum(!is.na(d2$pco2_fb)) # total non-NA pCO2 values
count_phINT_no_NA <-  sum(!is.na(d2$phINT)) # total non-NA pH seafet values
# calculate the distribution of non-NA pCO2 and pH values per month
d2_summary <- d2 %>% 
  dplyr::group_by(montht) %>% 
  dplyr::summarise(pco2_pc = round(100 * sum(!is.na(pco2_fb))/count_pco2_fb_no_NA, digits = 1),
                   pH_pc = round(100 * sum(!is.na(phINT))/count_phINT_no_NA, digits = 1)
                   )
# plot the distribution of non-NA pCO2 and pH values per month
d2_summary_long <- d2_summary %>% 
  tidyr::pivot_longer(cols = c(pco2_pc, pH_pc))
p_avail_month <- ggplot(data = d2_summary_long) +
  geom_col(aes(x = montht, y = value, fill = name), position = "dodge", width = 0.75) +
  scale_fill_discrete(name = "", labels = c("pCO2", "pH")) + 
  labs(x="Month", y="Percent of total number of data", title = "") +
  Mytheme(size_labs = 8) +
  theme(plot.margin = margin(t = 0, r = 0, b = 0, l = 0.5, unit = "cm"))
#ggsave(filename = "figures/monthly_availability.png", plot = p_avail_month, width = 15, height = 8, units = "cm")

@@

```{r whole time series, fig.width=10, fig.height=10, echo=FALSE, message= FALSE}
size_point <- 2 # size of data points
alpha <- 1 #transparency

d <- d_all %>% 
  as_tsibble(index = datetime) %>% 
  dplyr::arrange(datetime) %>%
  tsibble::fill_gaps() %>%
  dplyr::mutate(year_month = format(datetime, '%Y-%m'),
                quarter = quarter(datetime)) %>%
#  dplyr::rename(pco2_contros = pco2_corr) %>% 
  replace(., is.na(.), NA) # replace NaN by NA everywhere

d_ts <- d %>%
  dplyr::select(datetime, temp_insitu_11m, sal_fb, par_air, pco2_fb,
                pHint_tot_sf_9m, ph_s_sf_temp_insi, at_calc, at) #%>%
  #tidyr::pivot_longer(-datetime, values_to = "value")

#sal_fb
ts_sal <- d_all %>%
  ggplot(aes(x = datetime, y = sal_fb), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="Salinity FerryBox",x="", y="Salinity") +
  Mytheme(size_labs = 8)
# ts_sal_py <- ggplotly(ts_sal,dynamicTicks = TRUE)
# filename.html <- paste0(path_fig, "ts_sal_py.html")
# saveWidget(ts_sal_py, filename.html, selfcontained = T, libdir = "lib")

# temp_insitu_11m
ts_temp <- d_all %>%
  ggplot(aes(x = datetime, y = temp_insitu_11m), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="In situ temperature 11 m",x="", y="Temperature (°C)") +
  Mytheme(size_labs = 8)

# PAR
ts_par_air <- d_all %>%
  ggplot(aes(x = datetime, y = par_air), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="Atmospheric PAR",x="", y="PAR (umol m-2 s-1)") +
  Mytheme(size_labs = 8)

# PCO2 
ts_pCO2 <- d_all %>%
  ggplot() +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(aes(x = datetime, y = pco2_fb), 
             colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  geom_point(aes(x = datetime, y = pco2_calc_fb), colour="red", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="Mesured (blue) and calculated (CT+AT; red) pCO2",x="", y="pCO2 (uatm)") +
  Mytheme(size_labs = 8)

# SEAFET pH
ts_pH <- d_all %>%
  ggplot(aes(x = datetime, y = pHint_tot_sf_9m), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  geom_point(aes(x = datetime, y = ph_s_sf_temp_insi), colour="red", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="In situ pH below 8 m (seaFET in blue, spectrophotometric pH in red))",x="", y=expression(paste(pH[T]))) +
  Mytheme(size_labs = 8)

# AT
ts_at <- d_all %>%
  ggplot(aes(x = datetime, y = at_calc), na.rm=TRUE) +
  scale_x_datetime(breaks = date_breaks("1 year"), labels = date_format("%Y")) +
  geom_point(colour="blue", alpha=alpha, na.rm=TRUE, size=size_point) + 
  geom_point(aes(x = datetime, y = at), colour="red", alpha=alpha, na.rm=TRUE, size=size_point) + 
  labs(title="Total alkalinity calculated from salinity (blue) and discrete measurements (red)", x="", y="AT (umol/kg)") +
  Mytheme(size_labs = 8)

# Combine plots
g <- (ts_sal / ts_temp / ts_par_air / ts_pCO2 / ts_pH / ts_at)
g <- g + plot_annotation(tag_levels = 'A')
g

g <- cowplot::plot_grid(ts_sal, ts_temp, ts_par_air, ts_pCO2, ts_pH, ts_at, align='v', ncol=1)
print(g)
```
## Maintenance
<div style="text-align: justify">
